---
title: "Section1"
author: "Or Caspi"
date: "10/8/2020"
output: html_document
---

### Get API data:
```{r}
library(RSocrata)

url <- "https://data.austintexas.gov/resource/7d8e-dm7r.json"
df <- read.socrata(url)
```

Pre-cleaning:
-less than .1 mile trip distnace.
-more than 500 mile trip distance.
-more than 24hr trip drip duration.

### Scooters only:
```{r}
scooters <- df[which(df$vehicle_type=="scooter"),]
```

### Additional cleaning:
```{r}
# If exceeds 50 km/h:
scooters$trip_distance <- as.integer(scooters$trip_distance)
scooters$trip_duration <- as.integer(scooters$trip_duration)
scooters$speed <- scooters$trip_distance/scooters$trip_duration*3.6 # (m/s -> km/h)
scooters <- scooters[c(which(scooters$speed<=50)),]
N <- nrow(df)
n <- nrow(scooters)
print(paste("Droped",N-n,"trips"))
```

### Setting dates:
```{r}
scooters$Date <- as.character(scooters$start_time)
scooters$Date <- substr(scooters$Date, 1, 10)
scooters$Date <- as.Date(scooters$Date, "%Y-%m-%d")
```

### Getting COVID Cases data:
```{r}
urlc <- "https://raw.githubusercontent.com/caspior/DIC1/main/Austin_Covid.csv"
covid <- read.csv(urlc)
covid$Date <- as.Date(covid$Date, "%m/%d/%Y")
```

### Limit scooters dataset to covide period:
```{r}
sc <- scooters[which((scooters$Date >= min(covid$Date) & (scooters$Date <= max(covid$Date)))),]
```

### Adding daily cases to trip data set:
```{r}
sc <- merge.data.frame(sc,covid, by = "Date")
```

### Trips per day
```{r}
library(dplyr)

by_day <- group_by(sc, Date)
by_day <- summarise(by_day, trips = n(), avg_distance = mean(trip_distance), avg_duration = mean(trip_duration), sum_duration = sum(trip_duration), cases = min(Cases))
```

## Descriptive statistics

### Cases - Trips
```{r}
library(ggplot2)
colors <- c("Trips" = "blue", "Cases" = "red")

ggplot(by_day, aes(x=Date)) +
    geom_line(aes(y=cases, color="Cases"), size=2) +
    geom_line(aes(y=trips, color="Trips")) +
    labs(title = "Trips and COVID19 cases per day", x="Date", y="Count", color="Legend") +
    scale_x_date(date_breaks="1 month",date_labels="%b %Y") +
    scale_color_manual(values = colors) +
    theme_bw()
```

```{r}
ggplot(by_day, aes(x=Date)) +
    geom_line(aes(y=cases/max(cases), color="Cases"),size=2) +
    geom_line(aes(y=trips/max(trips), color="Trips"),size=0.7) +
    labs(title = "Trips and COVID19 cases per day", x="Date", y="Ratio of maximum", color="Legend") +
    scale_x_date(date_breaks="1 month",date_labels="%b %Y") +
    scale_color_manual(values = colors) +
    theme_bw()
```

```{r}
lm <- lm(by_day$cases~by_day$trips)

ggplot() +
    geom_point(data=by_day,aes(x=cases, y=trips),size=0.7) +
    geom_smooth(method="lm", data=by_day, aes(x=cases, y=trips), color="blue",size=0.7, ) +
    annotate("text", x=max(by_day$cases)/4, y=max(by_day$trips)*0.75, label=paste0("Trips = ",round(lm[["coefficients"]][["(Intercept)"]],2)," + ",round(lm[["coefficients"]][["by_day$trips"]],2)," * Cases")) +
    labs(title = "Trips as result of COVID19 cases per day", x="Cases", y="Trips") +
    theme_bw()
```

### Cases - Average Distance
```{r}
colors <- c("Average Distance" = "blue", "Cases" = "red")

ggplot(by_day, aes(x=Date)) +
    geom_line(aes(y=cases, color="Cases"), size=2) +
    geom_line(aes(y=avg_distance, color="Average Distance")) +
    labs(title = "Average distance and COVID19 cases per day", x="Date", y="Count", color="Legend") +
    scale_x_date(date_breaks="1 month",date_labels="%b %Y") +
    scale_color_manual(values = colors) +
    theme_bw()
```

```{r}
ggplot(by_day, aes(x=Date)) +
    geom_line(aes(y=cases/max(cases), color="Cases"),size=2) +
    geom_line(aes(y=avg_distance/max(avg_distance), color="Average Distance"),size=0.7) +
    labs(title = "Average distance and COVID19 cases per day", x="Date", y="Ratio of maximum", color="Legend") +
    scale_x_date(date_breaks="1 month",date_labels="%b %Y") +
    scale_color_manual(values = colors) +
    theme_bw()
```

```{r}
lm <- lm(by_day$cases~by_day$avg_distance)

ggplot() +
    geom_point(data=by_day,aes(x=cases, y=avg_distance),size=0.7) +
    geom_smooth(method="lm", data=by_day, aes(x=cases, y=avg_distance), color="blue",size=0.7, ) +
    annotate("text", x=max(by_day$cases)*0.75, y=max(by_day$avg_distance)*0.75, label=paste0("Average Distance = ",round(lm[["coefficients"]][["(Intercept)"]],2)," ",round(lm[["coefficients"]][["by_day$avg_distance"]],2)," * Cases")) +
    labs(title = "Average Distance as result of COVID19 cases per day", x="Cases", y="Average Distance") +
    theme_bw()
```

### Cases - Average Duration
```{r}
colors <- c("Average Duration" = "blue", "Cases" = "red")

ggplot(by_day, aes(x=Date)) +
    geom_line(aes(y=cases, color="Cases"), size=2) +
    geom_line(aes(y=avg_duration, color="Average Duration")) +
    labs(title = "Average duration and COVID19 cases per day", x="Date", y="Count", color="Legend") +
    scale_x_date(date_breaks="1 month",date_labels="%b %Y") +
    scale_color_manual(values = colors) +
    theme_bw()
    
```

```{r}
ggplot(by_day, aes(x=Date)) +
    geom_line(aes(y=cases/max(cases), color="Cases"),size=2) +
    geom_line(aes(y=avg_duration/max(avg_duration), color="Average Duration"),size=0.7) +
    labs(title = "Average duration and COVID19 cases per day", x="Date", y="Ratio of maximum", color="Legend") +
    scale_x_date(date_breaks="1 month",date_labels="%b %Y") +
    scale_color_manual(values = colors) +
    theme_bw()
```

```{r}
lm <- lm(by_day$cases~by_day$avg_duration)

ggplot() +
    geom_point(data=by_day,aes(x=cases, y=avg_duration),size=0.7) +
    geom_smooth(method="lm", data=by_day, aes(x=cases, y=avg_duration), color="blue",size=0.7, ) +
    annotate("text", x=max(by_day$cases)*0.75, y=max(by_day$avg_duration)*0.90, label=paste0("Average Duration = ",round(lm[["coefficients"]][["(Intercept)"]],2)," + ",round(lm[["coefficients"]][["by_day$avg_duration"]],2)," * Cases")) +
    labs(title = "Average Duration as result of COVID19 cases per day", x="Cases", y="Average Duration") +
    theme_bw()
```

### Cases - Total Duration
```{r}
colors <- c("Total Duration" = "blue", "Cases" = "red")

ggplot(by_day, aes(x=Date)) +
    geom_line(aes(y=cases, color="Cases"), size=2) +
    geom_line(aes(y=sum_duration/3600, color="Total Duration")) +
    labs(title = "sum_duration and COVID19-cases per day", x="Date", y="Daily cases") +
    scale_x_date(date_breaks="1 month",date_labels="%b %Y") +
    labs(title = "Total Duration and COVID19 cases per day", x="Cases", y="Count", color="Legend") +
    scale_color_manual(values = colors) +
    theme_bw()
```

```{r}
ggplot(by_day, aes(x=Date)) +
    geom_line(aes(y=cases/max(cases), color="Cases"),size=2) +
    geom_line(aes(y=sum_duration/max(sum_duration), color="Total Duration"),size=0.7) +
    labs(title = "Total duration and COVID19 cases per day", x="Date", y="Ratio of maximum", color="Legend") +
    scale_x_date(date_breaks="1 month",date_labels="%b %Y") +
    scale_color_manual(values = colors) +
    theme_bw()
```

```{r}
lm <- lm(by_day$cases~by_day$sum_duration)

ggplot() +
    geom_point(data=by_day,aes(x=cases, y=sum_duration),size=0.7) +
    geom_smooth(method="lm", data=by_day, aes(x=cases, y=sum_duration), color="blue",size=0.7, ) +
    annotate("text", x=max(by_day$cases)*0.75, y=max(by_day$sum_duration)*0.75, label=paste0("Total Duration = ",round(lm[["coefficients"]][["(Intercept)"]],2)," + ",round(lm[["coefficients"]][["by_day$sum_duration"]],4)," * Cases")) +
    labs(title = "Total Duration as result of COVID19 cases per day", x="Cases", y="Total Duration") +
    theme_bw()
```

# Trip patterns change

```{r}
library(tigris)
library(leaflet)
library(tmap)

austin <- tracts("TX", "Travis", year="2019")
```
```{r}
sc$month <- as.integer(sc$month)
by_start <- group_by(sc[which(sc$month<8),], census_geoid_start)
by_start <- summarise(by_start, trips = n())
by_end <- group_by(sc[which(sc$month>7),], census_geoid_start)
by_end <- summarise(by_end, trips = n())
by_end <- merge.data.frame(by_end, by_start, by = "census_geoid_start")
by_end$diff <- by_end$trips.x - by_end$trips.y
```
```{r}
departures <- geo_join(austin, by_end, "GEOID", "census_geoid_start")
```
```{r}
tm_shape(departures) +
  tm_fill("diff", style="quantile", n=7, palette="Greens", title="") +
  tm_layout(title = "Increase in shared scooter departures \n  in Austin, Texas due to the pandemic",title.position = c(.35, .93), title.size = 1.1)
```